Sensitivity Analysis

### 2024-01-12 SFE_article_Sensitivity_Analysis/Article_Model_Parameters.tex
12.01

\section{Introduction}
%\subfile{Sections/introduction}
\subfile{Sections/introduction_imp}

>>>>

\section{Introduction}
%\subfile{Sections/introduction}
\subfile{Sections/introduction_imp}
This study aims to analyse the influence of the operating conditions on the supercritical extraction process described in {\color{red}article 1}. To achieve this, The local sensitivity analysis methods is applied to obtain the sensitivity equations. The influence of mass-flow rate, pressure and the inlet temperature are considered.

\subfile{Sections/Literature_Review}

The study is structured as follows: Chapter \ref{CH: Thermodynamic} provides a general discussion on supercritical fluids to familiarize the reader with their properties. Chapter \ref{CH:Governing_equations_chapter} introduces the general balance equations. The balance equations are combined with the extraction kinetic equation to develop the process model in Chapter \ref{CH: Extraction_model}. The derivative-based local sensitivity analysis is presented in Chapter \ref{CH: Sensitivity_Analysis} and is then combined with the process model. Finally, the sensitivity plots are discussed in Chapters \ref{CH: Results} and \ref{CH: Conclusion}.

### 2024-01-12 SFE_article_Sensitivity_Analysis/Sections/Results_Sensitivity.tex
12.01

	The sensitivity equations were solved simultaneously with the original process model. This work investigates the influence of inlet temperature, pressure, and mass flow rate on the state space and the extraction yield. The process model and parameters have been discussed in {\color{red}article 1}. The process model was calibrated on set of experiments obtained at different operating conditions, $40^\circ C$ - $50^\circ C$ and 200 bar - 300 bar. The sensitivity analysis has been performed under the assumption that the system operates are $45^\circ C$, 250 bar, 0.4 l/min. The simulation was performed with the time step of 0.01 [min] and 400 equally distributed points in the spatial direction.

@@

    The flow rate change affects the fluid inside the system, but appropriate boundary conditions must be set. When utilizing Dirichlet boundary conditions, it's crucial to be aware of potential numerical discrepancies stemming from the differing methods used to compute them. The system's enthalpy is determined through the time evolution of governing equations, while the inlet's enthalpy depends on the inlet temperature and pressure. A minor numerical mismatch between these values may manifest as an enthalpy difference propagating along the spatial domain. To ensure consistency between the fluid at the inlet and inside the computational domain, this analysis employs Neumann boundary conditions.

@@

    Figure \ref{fig:Sensitivty_F_CF} illustrates how the concentration of solute in the fluid phase responds to an increase in the flow rate. Initially, sensitivities are close to zero, indicating a minimal system response. The growth in flow rate affects ${\color{blue}C_f}(z,t)$ by increasing the velocity, consequently elevating the concentration gradient. As a result, positive sensitivities emerge within the system, forming a front that progresses in the direction of flow. The positive front indicates that the larger amount of solute moves faster across the system, which leads to faster decrease of the total amount of solute in both phases. This reduced amount of solute in the solid phase restricts the extraction rate due to a diminishment of the concentration gradient. This limitation leads to the formation of a front composed of negative sensitivities, which advances through the extractor. The negative front indicates that the solute concentration in the fluid phase becomes lower than before the flow rate increment. Eventually, the negative sensitivities asymptotically approach zero.

@@

	As a result of the Low Mach-number assumption and the pressure deviation, temperature and density change simultaneously along the system. According to Equation \ref{EQ:Enthalpy_equation}, the pressure change directly affects the quantity $h \times \rho_f$ through $\partial ({\color{blue}P}(t) A_f) / \partial t$, leading to the step change presented in Figure \ref{fig:Sensitivty_P_H}. Depending on the configuration of the system, two cases are possible. Since the total energy in the extractor has changed, there may be a difference between the fluid inside the equipment and the boundary conditions at the inlet. If Dirichlet boundary conditions are applied, the inlet temperature is maintained at the predefined value and may differ from the temperature in the extractor. In such a case, the temperature difference will cause the heat front to propagate through the system. Alternatively, Neumann boundary conditions can be applied to ensure that the temperature inside the extractor matches that at its inlet. In this work, the second approach was chosen to simplify the discussion.

>>>>

	The sensitivity equations were solved simultaneously with the original process model. This work investigates the influence of inlet temperature, pressure, and mass flow rate on the state space and the extraction yield. The process model and parameters have been discussed in {\color{red}article 1}. The process model was calibrated on set of experiments obtained at different operating conditions, $40^\circ C$ - $50^\circ C$ and 200 bar - 300 bar. The sensitivity analysis has been performed under the assumption that the system operates are $45^\circ C$, 250 bar and 0.4 l/min.

@@

    The flow rate change affects the fluid inside the system, but appropriate boundary conditions must be set. When utilizing Dirichlet boundary conditions, it's crucial to be aware of potential numerical artifacts from the differing methods used to compute them. The system's enthalpy is determined through the time evolution of governing equations, while the inlet's enthalpy depends on the inlet temperature and pressure. A minor numerical mismatch between these values may manifest as an enthalpy difference propagating along the spatial domain. To ensure consistency between the fluid at the inlet and inside the computational domain, this analysis employs Neumann boundary conditions.

@@

    Figure \ref{fig:Sensitivty_F_CF} illustrates how the concentration of solute in the fluid phase responds to an increase in the flow rate. Initially, sensitivities are close to zero, indicating a minimal system response. The growth in flow rate affects ${\color{blue}C_f}(z,t)$ by increasing the velocity, consequently elevating the concentration gradient. As a result, positive sensitivities emerge within the system, forming a front that progresses in the direction of flow. The positive front indicates that the larger amount of solute moves faster across the system, which leads to faster decrease of the total amount of solute in both phases. This reduced amount of solute in the solid phase restricts the extraction rate due to a diminishment of the concentration gradient. This limitation leads to the formation of a front composed of negative sensitivities, which propagate through the extractor. The negative front indicates that the solute concentration in the fluid phase becomes lower than before the flow rate increment. Eventually, the negative sensitivities asymptotically approach zero.

@@

	As a result of the Low Mach-number assumption and the pressure deviation, temperature and density change simultaneously along the system. According to Equation \ref{EQ:Enthalpy_equation}, the pressure change directly affects the quantity $h \times \rho_f$ through $\partial ({\color{blue}P}(t) A_f) / \partial t$, leading to the step change presented in Figure \ref{fig:Sensitivty_P_H}. Depending on the configuration of the system, two cases are possible. Since the total energy in the extractor has changed, there may be a difference between the fluid inside the equipment and the boundary conditions at the inlet. If Dirichlet boundary conditions are applied, the inlet temperature is maintained at the predefined value and may differ from the temperature in the extractor. In such a case, the temperature difference will cause the heat front to propagate through the system. Alternatively, Neumann boundary conditions can be applied to ensure that the temperature inside the extractor matches that at its inlet. In this work, the second approach was chosen.

### 2024-01-12 SFE_article_Sensitivity_Analysis/Sections/introduction_imp.tex
12.01

	The economic feasibility of the process is crucial when selecting the appropriate technology. Conventional processes, such as distillation and organic solvent extraction, are frequently used for essential oil extraction. The distillation process is carried out at a high temperature, which causes thermal degradation of thermolabile compounds—considering that alternative techniques like supercritical fluid extraction gained popularity for extraction. Supercritical carbon dioxide, in particular, is attractive due to its unique properties such as inflammable, non-toxic, low critical temperature and non-corrosive. Furthermore, its critical point is relatively low compared to other fluids, making it a suitable alternative to traditional extraction techniques. The supercritical fluids exhibit both gas- and liquid-like properties so that operating conditions can adjust pressure-depended dissolving power.

@@

	One model proposed by \citet{Reverchon1993} is the hot ball model, which is based on an analogy to heat transfer and describes an extraction process from solid particles containing small quantities of solute where solubility is not a limiting factor. Another model, the Broken-and-Intact Cell model, was presented by \citet{Sovova1994}. This model describes a system where the outer surfaces of particles have been mechanically interrupted, allowing easy access of solvent to the solute from the broken cells. In contrast, the solute from the intact cells is less accessible due to high mass transfer resistance.

@@

	This study aims to develop a process model for extracting natural substances from solid materials and liquids using supercritical fluids, specifically supercritical $CO_2$. To achieve this, the model parameters are estimated based on thermodynamic relations and a set of experiments conducted at various conditions. The maximum likelihood estimator is used to solve the parameter estimation problem, and the obtained parameters are subjected to regression to derive correlations. These correlations enable the process model to be generalized across a range of temperatures (40 to $50~^\circ C$) and pressures (200 to 300 bar).

	The study is structured as follows: Chapter \ref{CH: Thermodynamic} provides a general discussion on supercritical fluids to familiarize the reader with their properties. Chapter \ref{CH:Governing_equations_chapter} introduces the general balance equations. The balance equations are combined with the extraction kinetic equation to develop the process model in Chapter \ref{CH: Extraction_model}. The maximum likelihood technique is presented in Chapter \ref{CH: Parameter_estimation} and is then combined with the process model. Chapter \ref{CH: Experiments} describes the experimental work and the data collected from experiments are used to estimate missing parameters related to the extraction kinetic. Finally, the parameter estimations and simulation results are discussed in Chapters \ref{CH: Results} and \ref{CH: Conclusion}

>>>>

	The economic feasibility of the process is crucial when selecting the appropriate technology. Conventional processes, such as distillation and organic solvent extraction, are frequently used for essential oil extraction. The distillation process is carried out at a high temperature, which causes thermal degradation of thermolabile compounds—considering that alternative techniques like supercritical fluid extraction gained popularity for extraction. Supercritical carbon dioxide, in particular, is attractive due to its unique properties such as inflammable, non-toxic, low critical temperature and non-corrosive. Furthermore, its critical point is relatively low compared to other fluids, making it a suitable alternative to traditional extraction techniques. The supercritical fluids exhibit both gas- and liquid-like properties so that operating conditions can adjust the dissolving power.

@@

	One model proposed by \citet{Reverchon1993} is the hot ball model, which is based on an analogy to heat transfer and describes an extraction process from solid particles containing small quantities of solute where solubility is not a limiting factor.

	Another model, the Broken-and-Intact Cell model, was presented by \citet{Sovova1994}. This model describes a system where the outer surfaces of particles have been mechanically interrupted, allowing easy access of solvent to the solute from the broken cells. In contrast, the solute from the intact cells is less accessible due to high mass transfer resistance.

@@
